---
title: "Code Summary for CAD Project"
author: " FIT3164 Team 07: Danzel, Xinhao and Yiqiu"
date: "2020/8/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains all codes with comments written for **FIT3164 Team 07 Heart Disease Project** in 2020.
A useful reference of using R Markdown: <http://rmarkdown.rstudio.com>.

## Section 0: Install and load necessary libraries.
Package Installation:
```{r eval=FALSE}
# install.packages("")
```

Package Loading:
```{r}
# library()
```

## Section 1: Data Preparation and Understanding.

This section includes reading, understanding and pre-processing of the Z-Alizadeh Sani dataset.

Note that Z-Alizadeh Sani dataset is available from: <https://archive.ics.uci.edu/ml/machine-learning-databases/00412/>.

Clear the work space before run the code.
```{r}
rm(list = ls())
```

Read the csv-formatted dataset as a dataframe.
```{r}
ZAS_Original = read.csv("Z_Alizadeh_Sani_Dataset.csv")
ZAS = ZAS_Original
```

Show the first 5 lines of the dataset to see what it looks like.
```{r}
head(ZAS)
```

Check the number of rows (patients) and columns (features).
```{r}
dim(ZAS)
```

There is a useless feature Exertional CP, remove this feature from the dataset (the reason of saying Exertional CP is useless is that it has only one unique value "N" which doesn't help predicting CAD at all).
```{r}
ZAS = ZAS[, !(colnames(ZAS) == "Exertional.CP")]
```

Convert all categorical features with only "N" / "Y" values to numerical binary (that is, to convert "N" to 0 and "Y" to 1).
```{r}
for (feature_index in 1:ncol(ZAS)) {
  # Get the name of the feature.
  feature_name = colnames(ZAS)[feature_index]
  # For each feature.
  feature = ZAS[, feature_index]
  # Extract unique values of each feature (in vectors).
  unique_val = unlist(c(unique(feature)))
  # If the feature has only "N" and "Y" categorical values, convert "N" to
  # 0 and "Y" to 1:
  if ((length(unique_val) == 2) &
      ("N" %in% unique_val) &
      ("Y" %in% unique_val)) {
    ZAS[ZAS[, feature_name] == "N", feature_name] = 0
    ZAS[ZAS[, feature_name] == "Y", feature_name] = 1
  }
}
```

Convert values of Sex ("Male" / "Fmale") and Cath ("Normal" / "Cad") features to binary (that is to: for Sex, convert "Fmale" to 0 and "Male" to 1; for Cath, convert "Normal" to 1 and "Cad" to 0).
```{r}
# Convert Sex to binary. Male is 1 and Female is 0.
ZAS$Sex[ZAS$Sex == "Male"] = 1
ZAS$Sex[ZAS$Sex == "Fmale"] = 0

# Convert Cath to binary. Normal is 0 and Cad is 1.
ZAS$Cath[ZAS$Cath == "Normal"] = 0
ZAS$Cath[ZAS$Cath == "Cad"] = 1
```

Convert values of features  (Function Class, BBB, Region RWMA and VHD) with more than two categories to dummy.
```{r}
# Function Class.
ZAS$Function.Class = as.factor(ZAS$Function.Class)
contrasts(ZAS$Function.Class) = contr.treatment(4)

# Region RWMA.
ZAS$Region.RWMA = as.factor(ZAS$Region.RWMA)
contrasts(ZAS$Region.RWMA) = contr.treatment(5)

# BBB and VHD.
ZAS$BBB = as.factor(ZAS$BBB)
contrasts(ZAS$BBB) = contr.treatment(3)

# VHD.
ZAS$VHD = as.factor(ZAS$VHD)
contrasts(ZAS$VHD) = contr.treatment(4)
```

All categorical variables are in characters or converted to numeric, this is just for convenience of pre-processing. Now we want them to be factors.
```{r}
for (feature_index in 1:ncol(ZAS)) {
  # Get the name of the feature.
  feature_name = colnames(ZAS)[feature_index]
  # For each feature.
  feature = ZAS[, feature_index]
  # Extract unique values of each feature (in vectors).
  unique_val = unlist(c(unique(feature)))
  # If the feature has only 0 and 1 numerical values:
  if ((length(unique_val) == 2) &
      (0 %in% unique_val) &
      (1 %in% unique_val)) {
    # Convert the feature from "character" to "numeric" object type.
    ZAS[,feature_name] = as.factor(ZAS[,feature_name])
  }
}
```

Check the current type of each feature to ensure they are in expected: there should be 34 categorical features (which should be all factor typed) and 21 quantitative features (which should be either numeric or integer typed).
```{r}
types = sapply(ZAS, class)
print(paste0("The number of factor typed features: ",
             length(grep("factor", types))))
print(paste0("The number of numeric / integer typed features: ", 
             length(grep("numeric", types)) + length(grep("integer", types))))
print(types)
```

Split the dataset into train and test sets (in 70%-30% ratio).
```{r}
# Set the seed to make all teammates have the same random dataset.
set.seed(3164)
# Get 70% of the dataset to be training dataset.
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train1 = ZAS[training.rows,]
# The rest of rows are testing dataset.
ZAS.test1 = ZAS[-training.rows,]

# Set the seed to make all teammates have the same random dataset.
set.seed(070707)
# Get 70% of the dataset to be training dataset.
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train2 = ZAS[training.rows,]
# The rest of rows are testing dataset.
ZAS.test2 = ZAS[-training.rows,]

# Set the seed to make all teammates have the same random dataset.
set.seed(31643164)
# Get 70% of the dataset to be training dataset.
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train3 = ZAS[training.rows,]
# The rest of rows are testing dataset.
ZAS.test3 = ZAS[-training.rows,]

# Set the seed to make all teammates have the same random dataset.
set.seed(13653)
# Get 70% of the dataset to be training dataset.
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train4 = ZAS[training.rows,]
# The rest of rows are testing dataset.
ZAS.test4 = ZAS[-training.rows,]

# Set the seed to make all teammates have the same random dataset.
set.seed(1654221)
# Get 70% of the dataset to be training dataset.
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train5 = ZAS[training.rows,]
# The rest of rows are testing dataset.
ZAS.test5 = ZAS[-training.rows,]
```