---
title: "Code Summary for CAD Project"
author: " FIT3164 Team 07: Danzel, Xinhao and Yiqiu"
date: "2020/8/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains all codes with comments written for **FIT3164 Team 07 Heart Disease Project** in 2020.
A useful reference of using R Markdown: <http://rmarkdown.rstudio.com>.

## Section 0: Install and load necessary libraries.
Installation:
```{r eval=FALSE}
# install.packages("readxl")
```

Loading:
```{r}
library(readxl)
```

## Section 1: Data Preparation and Understanding.

This section includes reading, understanding and pre-processing of the Z-Alizadeh Sani dataset.

Note that Z-Alizadeh Sani dataset is available from: <https://archive.ics.uci.edu/ml/machine-learning-databases/00412/>.

Clear the work space before run the code.
```{r}
rm(list = ls())
```

A small and quick note (to be resolved):
```{r}
# Is is necessary to create a log file recording which of the features have been
# formatted / changed / removed in what kind of way?
```

Since *.csv* format is more code-friendly than *.xlsx* format, convert the type of the dataset:
```{r eval=FALSE}
write.csv(read_excel("Z_Alizadeh_Sani_Dataset.xlsx"), 
          'Z_Alizadeh_Sani_Dataset.csv', 
          row.names = FALSE)
```

Read the converted csv dataset as a dataframe, copy one .
```{r}
ZAS_Original = read.csv("Z_Alizadeh_Sani_Dataset.csv")
ZAS = ZAS_Original
```

Show the first 5 lines of the dataset to see what it looks like.
```{r}
head(ZAS)
```

Check the number of rows (patients) and columns (features).
```{r}
dim(ZAS)
```

There is a useless feature Exertional CP, remove this feature from the dataset (the reason of saying Exertional CP is useless is that it has only one unique value "N" which doesn't help predicting CAD at all).
```{r}
ZAS = ZAS[, !(colnames(ZAS) == "Exertional.CP")]
```

Convert all categorical features with only "N" / "Y" values to binary (that is, to convert "N" to 0 and "Y" to 1).
```{r}
for (feature_index in 1:ncol(ZAS)) {
  # Get the name of the feature.
  feature_name = colnames(ZAS)[feature_index]
  # For each feature.
  feature = ZAS[, feature_index]
  # Extract unique values of each feature (in vectors).
  unique_val = unlist(c(unique(feature)))
  # If the feature has only "N" and "Y" categorical values, convert "N" to
  # 0 and "Y" to 1:
  if ((length(unique_val) == 2) &
      ("N" %in% unique_val) &
      ("Y" %in% unique_val)) {
    ZAS[ZAS[, feature_name] == "N", feature_name] = "0"
    ZAS[ZAS[, feature_name] == "Y", feature_name] = "1"
  }
}
```

Convert values of Sex ("Male" / "Fmale") and Cath ("Normal" / "Cad") features to binary (that is to: for Sex, convert "Fmale" to 0 and "Male" to 1; for Cath, convert "Normal" to 0 and "Cad" to 1).
```{r}
# Convert Sex to binary.
ZAS$Sex[ZAS$Sex == "Fmale"] = "0"
ZAS$Sex[ZAS$Sex == "Male"] = "1"
# Convert Cath to binary.
ZAS$Cath[ZAS$Cath == "Normal"] = "0"
ZAS$Cath[ZAS$Cath == "Cad"] = "1"
```

Convert values of features  (Function Class, BBB, Region RWMA and VHD) with more than two categories to dummy.
```{r}
#Function Class
ZAS$Function.Class = as.factor(ZAS$Function.Class)
contrasts(ZAS$Function.Class) = contr.treatment(4)

#Region RWMA
#transfer the data type from integer to factor
ZAS$Region.RWMA = as.factor(ZAS$Region.RWMA)
contrasts(ZAS$Region.RWMA) = contr.treatment(5)

#BBB and VHD
contrasts(ZAS$BBB) = contr.treatment(3)

#VHD
contrasts(ZAS$VHD) = contr.treatment(4)
```


Split the dataset into train and test sets (in 70%-30% ratio).
```{r}
#set seed to make all teammates have the same random dataset
set.seed(3164)
#get 70% of the dataset to be training dataset
training.rows = sample(1:nrow(ZAS), 0.7*nrow(ZAS))
ZAS.train = ZAS[training.rows,]
#the rest of rows are testing dataset
ZAS.test = ZAS[-training.rows,]
```