---
title: "Code Summary for CAD Project"
author: " FIT3164 Team 07: Danzel, Xinhao and Yiqiu"
date: "2020/8/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains all codes with comments written for **FIT3164 Team 07 Heart Disease Project** in 2020.
A useful reference of using R Markdown: <http://rmarkdown.rstudio.com>.

## Section 1: Data Preparation and Understanding.

This section includes reading, understanding and pre-processing of the Z-Alizadeh Sani dataset.

Note that Z-Alizadeh Sani dataset is available from: <https://archive.ics.uci.edu/ml/machine-learning-databases/00412/>.

### Covert the dataset type:
Since *.csv* format is more code-friendly than *.xlsx* format, we first convert the type of the dataset:
```{r eval=FALSE}
write.csv(read_excel("Z_Alizadeh_Sani_Dataset.xlsx"), 
          'Z_Alizadeh_Sani_Dataset.csv', 
          row.names = FALSE)
```
Read the converted csv dataset as a dataframe.
```{r}
ZAS = read.csv("Z_Alizadeh_Sani_Dataset.csv")
```
